\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A Radial Time Projection Chamber for $\alpha$ detection in CLAS at JLab}

### 2017-06-06 eg6-letters/NIM/EG6_RTPC_NIM.tex

The CLAS RTPC is a 250~mm long cylinder of 158~mm diameter, leaving just enough 
room to fit pre-amplifiers between the RTPC outer shell and the solenoid. The 
electromagnetic field is directed perpendicularly to the beam direction, 
such that drifting electrons are pushed away from the beam line. These electrons 
are amplified by three layers of curved Gas Electron Multiplier 
(GEM)~\cite{gem_sauli} and detected by the readout system 
on the external shell of the detector as illustrated in the figure~\ref{fig:RTPC_1_4}. The CLAS 
RTPC is segmented into two halves with independent GEM amplification systems that 
cover about 80\% of the azimuthal angle.

We detail here the different regions shown in figure~\ref{fig:RTPC_1_4} starting 
from the beam line and towards larger radius:\\
\begin{itemize}
   \item First, the 6 atm helium-4 target extends along the beam line axis, it 
      is 284~mm long with a 3~mm radius wall made of a 27-$\mu$m-thick Kapton.
   \item The first gas gap covers a radial range from 3 mm to 20 mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions from
      M\o ller electrons scattered by the beam. This 
      region is surrounded by a 4 $\mu$m thick window made of aluminized Mylar 
      is electronically grounded.
   \item The second gap region extends between 20~mm and 30~mm and is filled with a 
      gas mixture of 80$\%$ neon and 20$\%$ dimethyl ether (DME). This region 
      is surrounded by a 4~$\mu$m thick window made of aluminized Mylar, which 
      serves as the cathode and is set at 4260~V.
   \item The drift region is filled with the same Ne-DME gas mixture and extends 
      from the cathode to the first GEM, 60 mm away 
      from the beam axis. The average electric field in this region is about 550~V/cm 
      and perpendicular to the beam.
   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66 mm. In this configuration, the first GEM layer 
      serves as the anode and each subsequent GEM is set at a lower voltage to
      obtain a strong ($\sim$1600 V/cm) electric field between the GEM foils. A 
      275~V bias is applied to each GEM for amplification.
   \item The readout board has an internal radius of 69~mm and collects charge
     from the GEMs. Preamplifiers are plugged directly on its outer side and
     transmit the signal to the data acquisition electronics.
\end{itemize}

\begin{figure}[tbp]
\centering
\includegraphics[scale=0.70]{fig/GEM_photo.jpg}
\caption{Picture of a GEM foil similar to the one used for our RTPC. {\color{red} Source?}} 
\label{fig:GEMs}
\end{figure}

@@

One of the important problematic to develop the radial TPC was to obtain a 
good support structure of the GEM foils, allowing an easy installation of the 
GEM in good conditions and with high precision. At the same time, we wanted to 
keep the material budget small in the forward region where we detect other 
particles in subsequent detectors. We successfully realized these prerequisites 
by using fiber glass rings glued to each ends of the GEM foils to form self 
supporting cylinders that could be installed independently in the RTPC after 
all gluing and soldering operations. The rigidity of the GEM foil is enough
for the structure to be self supporting and only the 
backward part of the cylinder was fixed to the main mechanical structure. This 
design only left a light fiberglass ring in the forward region, reducing to a 
minimum secondary interactions.

@@

The drift paths are adjusted using a set of identified elastic events. We generate 
a helium nucleus in a GEANT4 simulation~\cite{GEANT4} of our RTPC for each event
based on the measured electron kinematic. Then, we compare the simulated trajectory of 
the Helium nuclei to the hits measured in the chamber. Because of the 
magnetic field, the drift paths are not linear in the RTPC. So to perform the extraction, 
we make a first approximation with a linear dependence between the radius of 
emission and the time of detection, and then refine our result. As it happens, 
the curvature is minimal and this process converge already on the second 
iteration.

@@

The second is a coherent noise affecting about 25\% of the pre-amplifiers, where the signature is simultaneous hits in most of the 16 channels in a preamp group.   An event-based technique to identify and remove it was developed based on counting simultaneous hits in preamp group, and, if sufficiently large, perform a dynamic pedestal subtraction based on the average ADC of the channels' neighbors within the preamp group.

@@

We reported on the construction, operation and calibration of a small GEM based RTPC 
designed to measure helium-4 nuclei in a high rate environment. The operation
of the detector was successful and allowed to detect helium nuclei at a rate
of 3.1 kHz in coincidence with the detection of high energy electrons and 
photons in the CLAS spectrometer.

@@

\bibitem{gem_sauli}
   F. Sauli, GEM: A new concept for electron amplification in gas detectors, 
   Nucl. Instr. and Meth. A 386, 531 (1997).

>>>>

The CLAS RTPC is a 250~mm long cylinder of 158~mm diameter, leaving just enough 
room to fit pre-amplifiers between the RTPC outer shell and the solenoid. The 
electromagnetic field is directed perpendicularly to the beam direction, 
such that drifting electrons are pushed away from the beam line. These electrons 
are amplified by three layers of curved GEMs and detected by the readout system 
on the external shell of the detector as illustrated in the 
figure~\ref{fig:RTPC_1_4}. The CLAS RTPC is segmented into two halves with 
independent GEM amplification systems that cover about 80\% of the azimuthal 
angle.

We detail here the different regions shown in figure~\ref{fig:RTPC_1_4} starting 
from the beam line and towards larger radius:\\
\begin{itemize}
   \item First, the 6 atm helium-4 target extends along the beam line axis, it 
      is 284~mm long with a 3~mm radius wall made of a 27-$\mu$m-thick Kapton.
   \item The first gas gap covers a radial range from 3 mm to 20 mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions from
      M\o ller electrons scattered by the beam. This 
      region is surrounded by a 4 $\mu$m thick window made of aluminized Mylar 
      is electronically grounded.
   \item The second gap region extends between 20~mm and 30~mm and is filled with a 
      gas mixture of 80$\%$ neon and 20$\%$ dimethyl ether (DME). This region 
      is surrounded by a 4~$\mu$m thick window made of aluminized Mylar, which 
      serves as the cathode and is set at 4260~V.
   \item The drift region is filled with the same Ne-DME gas mixture and extends 
      from the cathode to the first GEM, 60 mm away 
      from the beam axis. The average electric field in this region is about 550~V/cm 
      and perpendicular to the beam.
   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66 mm. In this configuration, the first GEM layer 
      serves as the anode and each subsequent GEM is set at a lower voltage to
      obtain a strong ($\sim$1600 V/cm) electric field between the GEM foils. A 
      275~V bias is applied to each GEM for amplification.
   \item The readout board has an internal radius of 69~mm and collects charge
     from the GEMs. Pre-amplifiers are plugged directly on its outer side and
     transmit the signal to the data acquisition electronics.
\end{itemize}

\begin{figure}[tbp]
\centering
\includegraphics[scale=0.70]{fig/GEM_photo.jpg}
\caption{Image of a typical GEM foil similar to the one used for our RTPC 
\cite{GEM_ref_pic}.}

   \label{fig:GEMs}
\end{figure}

@@

One of the important problematic to develop the radial TPC was to obtain a 
good support structure of the GEM foils, allowing an easy installation of the 
GEM in good conditions and with high precision. At the same time, we wanted to 
keep the material budget small in the forward region where we detect other 
particles in subsequent detectors. We successfully realized these prerequisites 
by using fiber glass rings glued to each ends of the GEM foils to form self 
supporting cylinders that could be installed independently in the RTPC after 
all gluing and soldering operations. The rigidity of the GEM foil is enough
for the structure to be self-supporting and only the backward part of the 
cylinder was fixed to the main mechanical structure. This design only left a 
light fiberglass ring in the forward region, reducing to a minimum secondary 
interactions.

@@

The drift paths are adjusted using a set of identified elastic events from our 
lower beam energy run. Based on the kinematic of the electrons in these events, 
we generate the helium nucleus in a GEANT4 simulation~\cite{GEANT4} of our 
RTPC. Then we compare the calculated GEANT4 trajectory of the Helium nuclei to 
the hits measured in the chamber.

Because of the magnetic field, the drift paths are not linear in the RTPC. So 
to perform the extraction, we make a first approximation with a linear 
dependence between the radius of emission and the time of detection, and then 
refine our result. As it happens, the curvature is minimal and this process 
converge already on the second iteration.

@@

The second is a coherent noise affecting about 25\% of the preamplifiers, where 
the signature is simultaneous hits in most of the 16 channels in a preamp 
group.   An event-based technique to identify and remove it was developed based 
on counting simultaneous hits in preamp group, and, if sufficiently large, 
perform a dynamic pedestal subtraction based on the average ADC of the 
channels' neighbors within the preamp group.

@@

We reported on the construction, operation and calibration of a small RTPC 
designed to measure helium-4 nuclei in high rate environment. The operation
of the detector was successful and allowed to detect helium nuclei at a rate
of 3.1 kHz in coincidence with the detection of high energy electrons and 
photons in the CLAS spectrometer.

@@

\bibitem{GEM_ref_pic}
   F. Sauli, The Gas Electron Multiplier (GEM): Operating Principles and 
   Applications, Nucl. Instr. and Meth, A805(2016)2
